Robust Bayesian Analysis of Causal Inference Problems

### 2024-03-28 JASA-template.tex
Small fixes (Matthias).

%TB to MT: did some modification in the following text
In particular through predictor selection, we also want 
to check if a certain bio-marker should be considered
for treatment decision. That is, in a very intuitive 
situation, we can observe blood pressure with some
other bio-markers and want to decide whether our
treatment guideline should also consider the blood
pressure of the subject before treating them. This is
useful as an unnecessary treatment of a subject can have severe consequences because of the medicinal
side effects. In general, this cannot be associated 
with a suitable loss function. Instead, we
assume that we can always revert any initial mistreatment by further treatments, 
and we can associate a loss function with the cost of further treatments.
So, we will associate two constant loss values $\ell_1$ and $\ell_2$ 
with false positives (falsely selected predictor) and 
false negatives (falsely rejected predictor) respectively. 
Clearly, false positives will lead to unwanted side effects and
false negatives will lead to mistreatment of the patient. Finally, we associate
a loss value $\ell_3$ for abstention from selecting
a variable which can be interpreted as the cost of further tests to determine whether that bio-marker
is important for constructing the treatment guideline.
Ideally, in most cases, $\ell_3\ll \ell_1,\ell_2$. However, in certain scenarios,
this might not be the case, especially when the condition of a subject deteriorates rapidly
over time.

Now, based on this notion of abstaining from selecting a variable, we can perform
a sensitivity analysis over a set of priors on the prior selection probability.
That is, we can consider a set of possible values for $q$ such that
$q\in\mathcal{P}$, where $\mathcal{P} \subseteq \left(0, 1\right)^{p}$.
Here, the equality occurs for the near vacuous case. However, in real-life
situations, performing a robust Bayesian analysis for the near vacuous case is 
not practical. Instead, we incorporate expert elicitation to define our model.
For instance, we can consider $q\in \left[\underline{q}, \overline{q}\right]$
where $p\underline{q}$ and $p\overline{q}$ represent the bounds of the prior expectation on the
total number of variables present in either of the models.

@@

The first two cases that we consider are:
\begin{description}
    \item[Scenario 1] --- $|\gamma_j|, |\beta_j|>0$ for $j\le 10$
    \item[Scenario 2] --- $|\gamma_j|>0$ for $j\le 10$ and $|\beta_j|>0$ for $j\le 15$
\end{description}
For both
cases, we consider different numbers of observations $n$ where
$n=25+ 5k$ for $k=0,1,2,\dots,35$ and $p=50$
predictors. This way, we check the efficiency of our method with varying level of information.

For the next two cases, we check our method for varying number of predictors (and hence sparsity level, i.e.\ the percentage of active variables present in the model):
\begin{description}
    \item[Scenario 3] --- $|\gamma_j|, |\beta_j|>0$ for $j\le 10$
    \item[Scenario 4] --- $|\gamma_j|>0$ for $j\le 10$ and $|\beta_j|>0$ for $j\le 15$
\end{description}
For both
cases, we consider different numbers of predictors $p$ where
$p=25+ 5k$ for $k=1,2,\dots,35$ and $n=100$
subjects.

@@

In these simulations, our method tends
to underestimate the causal effect.
%%% MT to TB: why is the problem with a? can you explain a bit more?
This suggests that
we may want to have a different value of $a$ for these sets of observations
%%% MT to TB: is the problem that it is "fixed"? this seems a bit misleading
%TB to MT: I was thinking of a differernt value to get a flatter prior but that felt like manipulation so kept it fixed for each sample size or predicrtor size
instead of a fixed value of $a=10$ for all of our analyses.
We can also see that the lower bound tends to improve
with increasing number of observations: as we accumulate
more information, the interval becomes smaller and converges towards
the true value.

@@

We show the result of our analyses with different number of predictors and
fixed number of sample in \cref{fig:comp:trt:pred,tab:causal2}. Similar to
the first two settings, we notice that our method is in good agreement with
BSSL for the third setting. However, for the fourth setting, our method
underestimate the treatment effect significantly (5\% in the worst case). This
is somewhat in sync with the second setting and a possible reason behind this
could be the fact that in both the cases, the number of active $\beta_j$'s is
higher than $\gamma_j$'s. This therefore, tends to reduce the value of treatment
effect. A possible way to work around this issue can be using the `DSS' estimates
of the regression coefficient and adjust the value of $\beta_T$.

Besides this interesting result. We notice that all of the four methods performs
perfectly in variable selection giving no misspecified variables. Therefore,
we omit the table of variable selection for third and fourth scenarios.

@@

Currently, the paper proposes a robust Bayesian approach for causal effect estimation where
we rely on sampling strategies to obtain the posterior bounds as well as performing 
variable selection. In future, it will be interesting to derive inner approximation bounds
for the posterior estimates to reduce the computational cost. Moreover, for the sake of
illustration, we rely on simple loss functions
%TB to MT: edited the conclusion
associated with the predictor selection. However,
this paradigm can be used for a generalised
decision theoretic framework as well where based on
the predictor selection we can formulate the 
problem as a decision support for treating a subject.

>>>>

%TB to MT: did some modification in the following text
In particular, through predictor selection, we also want 
to check if a certain bio-marker should be considered
for the treatment decision. For example, say we can observe blood pressure with some
other bio-markers and want to decide whether our
treatment guideline should also consider the blood
pressure of the subject before treating them. This is
useful as an unnecessary treatment of a subject can have severe consequences because of the medicinal
side effects. In general, it is hard to associate such consequences 
with a suitable loss function. Instead, we
assume that we can always revert any initial mistreatment by further treatments, 
and we can associate a loss function with the cost of further treatments.
So, we will associate two constant loss values $\ell_1$ and $\ell_2$ 
with false positives (falsely selected predictor) and 
false negatives (falsely rejected predictor) respectively. 
%%% MT to TB: not considering a predictor may or may not lead to mistreatment, it just means you'll have less information, not that the treatment is bad... I've chanted "will" to "may"
Clearly, false positives may lead to unwanted side effects and
false negatives may lead to mistreatment of the patient. Finally, we associate
a loss value $\ell_3$ for abstention from selecting
a variable which can be interpreted as the cost of further tests to determine whether that bio-marker
is important for constructing the treatment guideline.
Ideally, in most cases, $\ell_3\ll \ell_1,\ell_2$. However, in certain scenarios,
this might not be the case, especially when the condition of a subject deteriorates rapidly
over time.

Now, based on this notion of abstaining from selecting a predictor, we can perform
a sensitivity analysis over a set of priors on the prior selection probability.
That is, we can consider a set of possible values for $q$ such that
$q\in\mathcal{P}$, where $\mathcal{P} \subseteq \left(0, 1\right)^{p}$.
Here, the equality occurs for the near vacuous case. However, in real-life
situations, performing a robust Bayesian analysis for the near vacuous case is 
not practical. Instead, we incorporate expert elicitation to define our model.
For instance, we can consider $q\in \left[\underline{q}, \overline{q}\right]$
where $p\underline{q}$ and $p\overline{q}$ represent the bounds of the prior expectation on the
total number of variables present in either of the models.

@@

These four cases can be split in two categories. In one category we consider the
case where we have increasing number of observations. In the first category we have two
sub cases, in the first case we consider all active variables to be confounders
and in the second case we consider some active variables which are only related to
the outcome model. 
\begin{description}
    \item[Scenario 1a] --- $|\gamma_j|, |\beta_j|>0$ for $j\le 10$
    \item[Scenario 1b] --- $|\gamma_j|>0$ for $j\le 10$ and $|\beta_j|>0$ for $j\le 15$
\end{description}
For both
scenarios, we consider different numbers of observations $n$ where
$n=25+ 5k$ for $k=0,1,2,\dots,35$ and $p=50$
predictors. This way, we check the efficiency of our method with varying level of information.

For the second category, we check our method for varying number of predictors 
(and hence sparsity level, i.e.\ the percentage of active variables present in the model).
Similar to the first category we also have two sub cases:
\begin{description}
    \item[Scenario 2a] --- $|\gamma_j|, |\beta_j|>0$ for $j\le 10$
    \item[Scenario 2b] --- $|\gamma_j|>0$ for $j\le 10$ and $|\beta_j|>0$ for $j\le 15$
\end{description}
For both
cases, we consider different numbers of predictors $p$ where
$p=25+ 5k$ for $k=1,2,\dots,35$ and $n=100$
subjects.

@@

In these simulations, our method tends
to underestimate the causal effect.
%%% MT to TB: why is the problem with a? can you explain a bit more?
This suggests that
we may want to have a different value of $a$ for these different simulated datasets
%%% MT to TB: is the problem that it is "fixed"? this seems a bit misleading
%TB to MT: I was thinking of a differernt value to get a flatter prior but that felt like manipulation so kept it fixed for each sample size or predicrtor size
instead of a fixed value of $a=10$ for all of our analyses.
We can also see that the lower bound tends to improve
with increased number of observations: as we accumulate
more information, the interval becomes smaller and converges towards
a single value.

@@

We show the result of our analyses with different number of predictors and
fixed number of sample in \cref{fig:comp:trt:pred,tab:causal2}. Similar to
the first two settings, we notice that our method is somewhat in agreement with
BSSL for the third setting. However, for the fourth setting, our method
underestimates the treatment effect significantly (5\% in the worst case). This
is somewhat in sync with the second setting, and a possible reason behind this
could be the fact that in both the cases, the number of active $\beta_j$'s is
higher than $\gamma_j$'s. This therefore, tends to reduce the value of treatment
effect. A possible way to work around this issue can be using the `DSS' estimates
of the regression coefficient and adjust the value of $\beta_T$.

Besides this interesting result, we notice that all of the four methods perform
perfectly in variable selection giving no misspecified variables. Therefore,
we omit the table of variable selection for third and fourth scenarios.

@@

Currently, the paper proposes a robust Bayesian approach for causal effect estimation where
we rely on sampling strategies to obtain the posterior bounds as well as performing 
variable selection. In future, it will be interesting to derive inner approximation bounds
for the posterior estimates to reduce the computational cost. Moreover, for the sake of
illustration, we rely on simple loss functions
%TB to MT: edited the conclusion
associated with the predictor selection. However,
this paradigm can be used for a generalised
decision theoretic framework as well where based on
the predictor selection we can formulate the 
problem as a decision support problem for treating a subject.
